# (APPENDIX) Appendices {-}

# Appendix A: Updated operating model components
\beginAppA
## Updated ageing error matrix

The Sablefish age-structured assessment model relies on catch-at-age 
data to estimate the true age-composition of the population; however, 
observed catch-at-age data are based on otolith readings that are 
imperfectly known. Failure to account for errors in otolith readings 
may lead to smoothing estimates of age-classes, making it more 
difficult to detect strong recruitment years or stock-recruit relationships 
[@hanselman2012statistical]. Ageing errors may also bias estimates of 
growth parameters, maturity schedules, and natural mortality that 
can lead to overfishing or inaccurate yield projections 
[@lai1987effects; @tyler1989implications]

To account for ageing-error, the Sablefish age-structured operating
model uses an ageing error matrix. In this MSE cycle, we simplified the 
formulation of the ageing-error matrix from the previously 
used double-geometric model to a discretized normal distribution. The 
two major differences between these two formulations are (i) that the 
error structure is constrained 
to be symmetric for the normal formulation, while the double geometric 
model allows for some skew in the error distribution; and (ii) the
normal assumes  the assigned true age is the mode of the normal
density, forcing ageing errors to be on average unbiased.

We developed our ageing error matrix using otoliths that had been 
read by two different readers at the DFO Pacific Biological Station 
ageing lab. These data account for 
approximately 15\% of the total otolith readings for BC Sablefish, 
which are read first by the primary reader and then by a secondary 
reader as a quality control. In the majority of cases both readers 
agreed (62\%) and in cases where the two readings differ (38\%), both 
readers conferred to resolve the discrepancy and agree on the final 
age assigned (Pers. Comm, J. Groot, DFO). In most cases the final 
age reading was that assigned by the secondary or primary reader (36\%), 
but in a few cases a new age was assigned (2\%).

We applied statistical models for estimating the probability of observing 
an age class (a) given the true age (b) based on methods described in 
@richards1992statistical and @heifetz1999age. The model assumes a 
normal ageing-error distribution where the estimated standard deviation 
of the observed age for a true age b is based on three parameters 
$\Phi = \{ \sigma_1, \sigma_A, \alpha \}$ in the form:

\begin{equation}
\sigma(b) = \left\{
    \begin{array}{ll}
        \sigma_1 + (\sigma_A - \sigma_1) \frac{1 - e^{-\alpha(b - 1)} }{1 - e^{-\alpha(A - 1)}}, & \alpha \neq 0; \\
        \sigma_1 + (\sigma_A - \sigma_1) \frac{b-1}{A-1}, & \alpha = 0.\\
    \end{array} \right.
\end{equation}

Parameters $\sigma_1$ and $\sigma_A$ are the standard deviations for 
$b=1$ and $b=A$, representing the minimum and maximum ages, respectively. 
The $\alpha$ parameter determines the non-linearity of the function, such 
that~$\sigma(b)$ becomes linear as $\alpha \rightarrow 0$. The age-error matrix 
is defined as:

\begin{align}
q(a ~|~ b, \Phi) &= \frac{x_{ab}(\Phi)}{\sum_{a = 1}^A x_{ab}(\Phi) }; \\
x_{ab} &= \frac{1}{\sqrt{2\pi}\sigma(b)} e^{-\frac12 \left[ \frac{a-b}{\sigma(b)} \right]^2}.
\end{align}

Given that the true age of the fish is unknown, it is not possible 
to accurately determine bias in the age readings and whether certain 
age classes are more likely to be under or over-estimated. We tested
2 different approaches for the assumed “true age”, using 1) the 
mean of the two reader ages rounded to the nearest integer 
[@heifetz1999age], and 2) the final age assigned. For both 
approaches we set $A=90$, based on the maximum assigned age by the 
readers. 

The likelihood $\mathcal{L}$ of observed ages $A$ given true ages B is then defined 
as:

\begin{equation}
\mathcal{L}(A|B) = \prod_{i = 1}^I \prod_{j = 1}^J q(a_{ij} ~|~ b_i \Phi),
\end{equation}

where $b_i$ is the assumed ‘true age’ of fish $i$, and $a_{ij}$ is 
the age assigned by reader $j$ to the individual fish $i$. Maximum 
likelihood parameter estimates, predicted standard deviation at age, and age-error 
matrices are provided below (Table A1, Fig. A1 & A2)


## Trawl Age-Length Key and updated selectivity curve

The Sablefish age-structured operating model uses observations
of catch at age from commercial fisheries to estimate natural 
mortality and gear selectivity functions. Trawl selectivity has
been identified a key determinant in reducing uncertainty in
estimates of sub-legal Sablefish catch and releases 
[@cox2019evaluating], as up until now the trawl selectivity
model was heavily dependent on priors for a normal selectivity 
curve estimated from tagged fish recovered (within one year
from release) in the commercial trawl fishery. To improve estimates 
of legal and sub-legal fishing mortality from the 
trawl sector, we leveraged catch-at-age and catch-at-length data
from BC trawl fisheries to develop a sex-specific age-length key, 
which was in turn used to increase the catch-at-age sample size.

To develop our age-length key, we used all available catch-at-age
data collected from observed trips in the commercial trawl fishery.
We then used this to populate an empirical age-length frequency 
matrix, binning fish into 3cm length bins and 1 year age classes.
We defined this matrix as
\begin{equation}
F = \left[ n_{l,a} \right],
\end{equation}
where $n_{l,a}$ is the number of fish observed in length bin $l$ 
and age class $a$. The matrix $A$ was converted to a probability 
of age-at-length $l$ matrix $P$ by normalising the columns of $A$
\begin{equation}
P_{l,a} = F_{l,a} / \sum_{a'} F_{l,a'}. 
\end{equation}

We then generated expected age composition data by applying the
matrix $P$ to length compositions $C_l$ derived from the
commercial trawl catch-at-length data.
\begin{align}
C_a &= P^T \cdot C_l,
\end{align}
where $P$ is transposed so that the length dimension matches
the vector $C_l$. We restricted $C_l$ to catch-at-length data
from years where at least 5 trips were sampled. We defined keys 
$P_m$ and $P_f$ for male and female fish, respectively, and 
generated sex-specific age observations (Figures A3 and A4).
Length observations from unsexed fish were treated as 
male specimens, as the operating model optimisation would
not converge when they were treated as females.

Inferred catch-at-age compositions had a noticable effect
on the selectivity-at-length curves for the trawl fleet (Figure A5). 
The fully selected size class moved from about 42 cm to 48 cm, and the 
shape of the Gamma selection curve dome was narrower, deselecting
to about 60% by the 55cm size limit, as opposed to about 80%
for the normal model in 2016.

\newpage


### Tables

```{r, echo = FALSE, include = FALSE, warning = FALSE}
ageErrTable <- tibble::tribble( 
    ~Case,   ~TrueAge,           ~sigma1, ~sigmaA, ~alpha, 
    1,      "Mean Reader Age",    0.38,    4.80,    0.014,
    2,      "Final Age Assigned", 0.89,    9.35,    -0.008, )

colnames(ageErrTable) <- c( "Case",
                            "True Age",
                            "$\\sigma_1$",
                            "$\\sigma_A$",
                            "$\\alpha$" )
```

```{r, echo = FALSE, warning = FALSE, results = "as-is"}
csasdown::csas_table(  ageErrTable,
                       caption = "Ageing error model parameters for both true age cases tested.",
                       booktabs = TRUE, escape = FALSE,
                       font_size = 12 ) %>%
    kable_styling(  latex_options = c("hold_position"),
                    bootstrap_options = c("striped", "hover","scale_down")     )
```

\newpage

### Figures
```{r, echo = FALSE, warning = FALSE, include = FALSE}

ageErr1Cap <- "Estimated standard deviation of observed ages for the two age assignment cases considered."
ageErr2Cap <- "Probability of observed ages given the true age indicated in top right corner of each panel for the two age assignment cases considered."


maleAgeCap <- "Inferred male catch-at-age compositions generated by the trawl age-length key from length observations of male and unsexed fish."
femaleAgeCap <- "Inferred female catch-at-age compositions generated by the trawl age-length key from length observations of female fish"


trawlSelCap <- "Trawl selectivity-at-length curves from the 2016 operating model (dashed grey line) and 2019 operating model (solid black line), and the legal size limit (vertical red dashed line). The length axis starts at the modeled length at age-1 of 32cm."
```


```{r, warning = FALSE, out.width = "90%", fig.align = "center", fig.cap=ageErr1Cap, echo = FALSE}
knitr::include_graphics("data/ageErr1")
```

\newpage

```{r, warning = FALSE, out.width = "90%", fig.align = "center", fig.cap=ageErr2Cap, echo = FALSE}
knitr::include_graphics("data/ageErr2")
```

\newpage

```{r, warning = FALSE, out.width = "90%", fig.align = "center", fig.cap=maleAgeCap, echo = FALSE}
knitr::include_graphics("data/base_ALK_mUnsexed/plotObsAgeFreq_m3")
```

\newpage

```{r, warning = FALSE, out.width = "90%", fig.align = "center", fig.cap=femaleAgeCap, echo = FALSE}
knitr::include_graphics("data/base_ALK_mUnsexed/plotObsAgeFreq_f3")
```

\newpage

```{r, warning = FALSE, out.width = "90%", fig.align = "center", fig.cap=trawlSelCap, echo = FALSE}
knitr::include_graphics("data/trawlSelOverlay")
```







